Spin effect on Christodoulou memory for merging blackhole binaries

### 2018-12-13 notes/Paper.tex
added notes on paper

\documentclass[prd,preprintnumbers,twocolumn,eqsecnum,floatfix,letter]{revtex4}
\usepackage{color}
\usepackage{calc}
\usepackage{amsmath,amssymb,graphicx}
\usepackage{amssymb,amsmath}
\usepackage{tensor}
\usepackage{bm}
\usepackage{times}
\usepackage[varg]{txfonts}
\usepackage[colorlinks, pdfborder={0 0 0}]{hyperref}
\definecolor{LinkColor}{rgb}{0.75, 0, 0}
\definecolor{CiteColor}{rgb}{0, 0.5, 0.5}
\definecolor{UrlColor}{rgb}{0, 0, 0.75}
\hypersetup{linkcolor=LinkColor}
\hypersetup{citecolor=CiteColor}
\hypersetup{urlcolor=UrlColor}
\maxdeadcycles=1000
\allowdisplaybreaks
\textwidth 7.5 in 
\hoffset -1 cm 
\newcommand{\comment}[1]{\textcolor{blue}{\textit{#1}}}
\newcommand{\ajith}[1]{\textcolor{red}{\textit{Ajith:#1}}}
\newcommand{\ashok}[1]{\textcolor{cyan}{\textit{Ashok:#1}}}

@@

\newcommand{\LIGO}{LIGO Laboratory, California Institute of Technology, 
Pasadena, CA 91125, USA}
\newcommand{\CIT}{Theoretical Astrophysics, California Institute of
Technology, Pasadena, CA 91125, USA}
\newcommand{\UWM}{UWM}

\title{Notes on the SpinTaylorF2 project}
\author{P.~Ajith}\email{ajith@icts.res.in}
\affiliation{International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bangalore 560012, India}
\author{Ashok Choudhary}
\affiliation{International Centre for Theoretical Sciences, Tata Institute of Fundamental Research, Bangalore 560012, India}

\begin{abstract}
\end{abstract}

@@

\section{Computing the spin and angular momentum vectors in the frequency domain}

\subsection{Performing the evolution in Cartesian coordinates}

The spin angular momentum vectors $\bS_i (v)$ and the unit vector $ \LNh (v)$ along the Newtonian angular momentum the frequency domain can be computed by using the evolution equations: 
\begin{eqnarray}
\frac{d \bS_i}{dv} & = & \frac{d \bS_i}{dt}  \frac{dt}{dv}, \\
\frac{d \LNh}{dv}  & = & \frac{d \LNh}{dt} \frac{dt}{dv},  
\label{eq:precessionEqFreqDom}
\end{eqnarray}
%
where 
\begin{equation}
\frac{d \LNh}{dt}  = - \frac{1}{||\mathbf{L}||} \, \frac{d}{dt} (\bS_1 + \bS_2),
\label{eq:orbitPrec}
\end{equation}
% 
%
\begin{equation}
\frac{d \bS_i}{dt}  =  \bm{\Omega}_i \times \bS_i \,,~ i = 1,2, 
\label{eq:spinPrec}
\end{equation}
%
and 
% 
\begin{widetext}
\begin{align}\label{eq:EbyF}
\frac{dt}{dv} & = \frac{5m}{32 \, v^9  \, \eta} \left\{1 + v^2 \left[\frac{11 \eta }{4}+\frac{743}{336} \right] 
    + v^3 \left[\frac{113}{12} \left(\biggl( 1-\frac{76\eta}{113}\biggr) \, \chisdL + \delta \, \chiadL \right)-4 \pi \right] \right. \nonumber \\ 
    &\quad + v^4 \left[(\chiadL)^2 \left(30 \eta -\frac{719}{96}\right)-\frac{719 \chiadL \, \chisdL \delta
   }{48}+\chiaSqr \left(\frac{233}{96}-10 \eta \right)+\frac{233 \chisDchia \delta}{48} \right. \nonumber\\
   &\qquad \left. +(\chisdL)^2 \left(-\frac{\eta }{24}-\frac{719}{96}\right)+\chisSqr \left(\frac{7 \eta
   }{24}+\frac{233}{96}\right)+\frac{617 \eta ^2}{144}+\frac{5429 \eta }{1008}+\frac{3058673}{1016064} \right] \nonumber \\ 
    &\quad + v^5 \left[\chiadL \delta  \left(\frac{7 \eta }{2}+\frac{146597}{2016}\right)+\chisdL \left(-\frac{17 \eta
   ^2}{2}-\frac{1213 \eta }{18}+\frac{146597}{2016}\right)+\frac{13 \pi  \eta }{8}-\frac{7729 \pi }{672} \right]  \nonumber\\ 
    &\quad + v^6 \left[\frac{1712 \gamma_E }{105}+\frac{25565 \eta ^3}{5184}-\frac{15211 \eta
   ^2}{6912}-\frac{451 \pi ^2 \eta }{48}+\frac{3147553127 \eta }{12192768}+\frac{32 \pi
   ^2}{3}-\frac{10817850546611}{93884313600}+\frac{1712 \ln (4v)}{105} \right]  \nonumber\\ 
    &\quad \left. + v^7 \left[\frac{14809 \pi  \eta ^2}{3024}-\frac{75703 \pi  \eta }{6048}-\frac{15419335 \pi }{1016064} \right] \right\}. 
\end{align}
\end{widetext}
% 
(\ajith{Need to add the BH absorption term above}). The (orbit averaged) precession frequencies $\bm{\Omega}_i$ appearing in Eq.~(\ref{eq:spinPrec}) are given by 
\begin{align}
\label{eq:Omega}
\bm{\Omega}_1 & =  \frac{v^5}{m} \left\{\left(\frac{3}{4}+\frac{\eta}{2}-\frac{3\delta}{4}\right) \LNh \right. \nonumber \\
     & \quad +  \frac{v}{2m^2} \left[-3 \, (\bS_2 + q \, \bS_1).\LNh ~ \LNh + \bS_2 \right]  \nonumber \\
     & \quad +  \left. v^2 \left(\frac{9}{16} + \frac{5\eta}{4} - \frac{\eta^2}{24} - \frac{9\delta}{16} + \frac{5\delta \eta}{8} \right)  \LNh  \right\}, \nonumber \\
%
\bm{\Omega}_2 & =  \frac{v^5}{m} \left\{\left(\frac{3}{4}+\frac{\eta}{2}+\frac{3\delta}{4}\right) \LNh \right. \nonumber \\
     & \quad +  \frac{v}{2m^2} \left[-3 \, (\bS_1 + q^{-1} \, \bS_2).\LNh ~ \LNh + \bS_1 \right]  \nonumber \\
     & \quad +  \left. v^2 \left(\frac{9}{16} + \frac{5\eta}{4} - \frac{\eta^2}{24} + \frac{9\delta}{16} - \frac{5\delta \eta}{8} \right)  \LNh  \right\}. 
\end{align}

%
Above, $v \equiv (m \omega)^{1/3}$ is the PN expansion parameter, $\omega$ is the orbital frequency, $m \equiv m_1 + m_2$ is the total mass of the binary, $q \equiv m_2/m_1$ is the mass ratio, $\eta \equiv m_1 m_2/m^2$ is the symmetric mass ratio, $\delta \equiv (m_1-m_2)/m$ the asymmetric mass ratio, $\gamma_E$ the Euler's constant, and $\LNh$ the unit vector along the Newtonian orbital angular momentum $\LN \equiv m^2\eta/v ~ \LNh$. The spin variables $\bchis$ and $\bchia$ are related to the (dimensionless) spin vectors of the binary as: 
%
\begin{equation}
\bchis = (\bchi_1 + \bchi_2)/2, ~~~ \bchia = (\bchi_1 - \bchi_2)/2,
\end{equation}
%
where $\bchi_i \equiv \bS_i/m_i^2$.

\subsection{Performing the evolution in spherical polar coordinates}

The time evolution equation of $i \equiv \arccos L_{Nz}$ and $ \alpha \equiv \arctan (L_{Ny}/L_{Nx})$ are given by:
\begin{equation}
\frac{d\iota}{dt} = \frac{-1}{\|L\| \sin\iota}\sum\limits_{k = 1}^{2}m^2_{k}\| \boldsymbol{\chi}_{k}\|\sin\theta_{s_{k}}\frac{d\theta_{s_{k}}}{dt}
\end{equation}  
and
\begin{widetext}
\begin{align}
\begin{split}
\frac{d\alpha}{dt} =& \frac{\cos\alpha}{\sin\iota}\Bigg(\frac{-1}{\|\boldsymbol{L}\|}\sum\limits_{k = 1}^{2} m^{2}_{k}\|\boldsymbol{{\chi}}_{k}\|\left[\frac{d\theta_{S_{k}}}{dt}\cos\theta_{s_{k}}\sin\phi_{s_{k}} + \frac{d\phi_{S_{k}}}{dt}\sin\theta_{s_{k}}\cos\phi_{s_{k}}\right]\Bigg) \\ 
- & \frac{\sin\alpha}{\sin\iota}\Bigg(\frac{-1}{\|\boldsymbol{L}\|}\sum\limits_{k = 1}^{2} m^{2}_{k}\|\boldsymbol{{\chi}}_{k}\|\left[\frac{d\theta_{S_{k}}}{dt}\cos\theta_{s_{k}}\cos\phi_{s_{k}} - \frac{d\phi_{S_{k}}}{dt}\sin\theta_{s_{k}}\sin\phi_{s_{k}}\right]\Bigg)
\end{split}
\end{align}
 \end{widetext}

 The angles corresponding to spin vectors($S_{1}$ and $S_{2}$)can also be written as:
 \begin{eqnarray}
 \frac{d\theta_{s}}{dt} = \frac{1}{\|\boldsymbol{\chi}\|\cos\theta_{s}}\left[
 \cos\phi_{s}(\Omega_{y}\chi_{z}-\Omega_{z}\chi_{y}) + \sin\phi_{s}(\Omega_{z}\chi_{x}-\Omega_{x}\chi_{z}) \right], \nonumber \\ 
  \frac{d\phi_{s}}{dt} = \frac{1}{\|\boldsymbol{\chi}\|\sin\theta_{s}}\left[
  \cos\phi_{s}(\Omega_{z}\chi_{x}-\Omega_{x}\chi_{z}) - \sin\phi_{s}(\Omega_{y}\chi_{z}-\Omega_{z}\chi_{y}) \right], \nonumber \\ 
i = 1,2 \end{eqnarray}
where the expression for $\Omega$ are given in \ref{eq:Omega}

\section{Computing the GW polarizations in the radiation frame}

In the absence of precession, the phase evolution of the (dominant harmonic) GWs is twice the orbital phase $\varphi(t)$. But the precession of the orbital plane introduces additional modulation in the orbital phase. If we define $\Phi(t)$ as the orbital phase with respect to the line of ascending nodes (the point at which the orbit crosses the $x-y$ plane from below) then the the phase evolution of the (dominant harmonic) GWs is given by $2\Phi(t)$, where $\Phi(t)$ is given by 
%
\begin{equation}
\label{eq:dPhiByDt}
\frac{d\Phi}{dt} = \omega - \frac{d\alpha}{dt} \cos i, 
\end{equation}
%
where $\alpha$ and $i$ are the angles describing the evolution of the orbital angular momentum vector $\LNh$ in the Finn-Chernoff coordinate system:
%
\begin{equation}
\alpha \equiv \arctan (\hat{L}_{Ny}/\hat{L}_{Nx}), ~~~ i \equiv \arccos (\hat{L}_{Nz}). 
\end{equation}
%
Time domain GW polarizations in the radiation frame can be written as 
%
\begin{eqnarray}
\label{eq:polzns_timedom}
h_{+,\times}(t) & = & \frac{-m \eta v^2}{d} ~\Bigg\{ C_{+,\times}\cos 2[\Phi(t)+\Phi_{0}] \nonumber \\ 
 & + & S_{+,\times}\sin 2[\Phi(t)+\Phi_{0}] \Bigg\}, \nonumber \\ 
\end{eqnarray} 
% 
which can be rewritten as 
%
\begin{eqnarray}
h_{+,\times}(t) = & A_{+,\times}(t) \, \cos \Psi_{+,\times}(t),
\end{eqnarray} 
%
where 
\begin{eqnarray}
\label{eq:polz_timedom_ampphase}
A_{+,\times}(t) & = & \frac{-m \eta v^2}{d} \sqrt{C_{+,\times}^2 + S_{+,\times}^2}, \nonumber \\ 
\Psi_{+,\times}(t) & = & 2 [\Phi(t) + \Phi_0] - \arctan \left(\frac{S_{+,\times}}{C_{+,\times}} \right),
\end{eqnarray}
where 
\begin{eqnarray}
C_+(t) & = & 2 \cos^2\alpha (\cos^2\Theta \cos^2 i+1)
	 -  \sin^2\alpha (\cos 2\Theta+\cos 2i+2) \nonumber \\ 
	& + & \cos \alpha \sin 2\Theta \sin 2i+2 \sin^2 \Theta \sin^2 i , \nonumber \\ 
S_+(t) & = & -2 \sin\alpha (\cos\alpha (\cos 2\Theta+3) \cos i + 2 \sin\Theta \cos\Theta \sin i), \nonumber \\ 
C_\times(t) & = & \sin \alpha (2 \cos \alpha \cos \Theta (\cos 2i+3)+4 \sin \Theta \sin i \cos i), \nonumber \\ 
S_\times(t) & = & 4 \cos 2\alpha \cos \Theta \cos i+ 4 \cos \alpha \sin \Theta \sin i.
\end{eqnarray}

\section{Computation of the Fourier transform}

In this section, we compute the Fourier transform of the time domain GW polarizations given in Eq.~(\ref{eq:polz_timedom_ampphase}).

\subsection{Stationary phase approximation}

We compute the Fourier transform of Eq.~(\ref{eq:polz_timedom_ampphase}) using stationary phase approximation (SPA). According to the SPA, the power in a Fourier frequency $f$ should mostly come from the saddle point $t_f$ given by the relation 
%
\begin{equation}
 f = F_{+,\times}(t_f).
 \label{eq:saddle_point}
\end{equation}
%
Above $F_{+,\times}(t_f)$ is the instantaneous GW frequency of the $+,\times$ polarization in the radiation frame. That is, 
\begin{equation}
\label{eq:saddle_point}
2 \pi F_{+,\times}(t) =  \frac{d\Psi_{+,\times}(t)}{dt} = 2 \omega_\mathrm{orb}(t) - \xi_{+,\times}(t), 
\end{equation}
where 
\begin{equation}
\xi_{+,\times}(t) = 2 \frac{d\alpha}{dt} \cos i - \frac{d}{dt} \arctan\left(\frac{S_{+,\times}}{C_{+,\times}}\right).
\end{equation}
% 
Then the Fourier transform can be written as as 
%
\begin{equation}
\tilde{h}_{+,\times}(f) = \mathcal{A}_{+,\times}(t_f) ~ e^{i[\Psi_{f+,\times}(t_f)-\pi/4]},
\end{equation}
where 
\begin{equation}
\label{eq:fourier_ampl}
\mathcal{A}_{+,\times}(t_f) = \frac{A_{+,\times}(t_f)}{2 \sqrt{\dot{F}_{+,\times}(t_f)}},
\end{equation}
and 
\begin{equation}
\label{eq:fourier_phase}
\Psi_{f\pc}(t_f) = 2\pi f t_{f}- 2 \varphi_\mathrm{orb}(t_f) + \int_{t_0}^{t_f} \xi_\pc(t) \, dt 
\end{equation}

The time-derivative of the instantaneous frequency appearing in Eq.~(\ref{eq:fourier_ampl}) can be written as 
%
\begin{eqnarray}
\frac{dF_\pc(t_f)}{dt} & = & \frac{dF_\pc}{dv} \frac{dv}{dt} \nonumber \\
 & = & -\frac{3v_f^2}{\pi m^2} \frac{\mathcal{F}(v_f)}{E'(v_f)}\left[1-\frac{m}{6v_f^2}\frac{d\xi_{+,\times}(t_f)}{dv} \right].
\end{eqnarray}
%
Using the binomial expansion of $(1-x)^{-1/2} \simeq (1+x/2)$, 
%
\begin{eqnarray}
\frac{1}{\sqrt{\dot{F}_{+,\times}(t_f)}} \simeq \sqrt{\frac{\pi}{3}} \frac{m}{v_f} \left[\frac{-E'(v_f)}{\mathcal{F}(v_f)} \right]^{1/2} \left[1+\frac{m}{12v_f^2} \frac{d\xi_{+,\times}(t_f)}{dv}\right]
\end{eqnarray}
%
Plugging this in Eq.~(\ref{eq:fourier_ampl}),
%
\begin{eqnarray}
\mathcal{A}_{+,\times}(t_f) & = &  \frac{m^2\eta v_f}{2d} \sqrt{\frac{\pi}{3}} \left[C_{+,\times}(t_f)^2 + S_{+,\times}(t_f)^2 \right]^{1/2} \nonumber \\ 
& & \left[\frac{-E'(v_f)}{\mathcal{F}(v_f)} \right]^{1/2} \left[1+ \frac{m}{12 v_f^2} \frac{d\xi_{+,\times}}{dv}\right]
\label{eq:spa_amp}
\end{eqnarray}
%
\subsection{Computing the saddle point}

For each Fourier frequency $f$, compute the corresponding saddle point $v_f$ (value of $v$ at the saddle point) by solving Eq.(\ref{eq:saddle_point}), which can be rewritten as 
%
\begin{equation}
v_f = v_f^0 \left[1+ \frac{\xi_{+,\times}(v_f)}{2\pi f} \right]^{1/3} \simeq  v_f^0 \left[1+ \frac{\xi_{+,\times}(v_f)}{6\pi f} \right]
\label{eq:saddle_point2}
\end{equation}
% 
where $v_f^0 = (\pi m f)^{1/3}$. Below, we describe the different approximations used in solving this equation.

>>>>

\documentclass[prd,preprintnumbers,twocolumn,eqsecnum,floatfix,letter]{revtex4}
\usepackage{color}
\usepackage{calc}
\usepackage{amsmath,amssymb,graphicx}
\usepackage{amssymb,amsmath}
\usepackage{tensor}
\usepackage{bm}
\usepackage{times}
\usepackage[varg]{txfonts}
\usepackage[colorlinks, pdfborder={0 0 0}]{hyperref}
\definecolor{LinkColor}{rgb}{0.75, 0, 0}
\definecolor{CiteColor}{rgb}{0, 0.5, 0.5}
\definecolor{UrlColor}{rgb}{0, 0, 0.75}
\hypersetup{linkcolor=LinkColor}
\hypersetup{citecolor=CiteColor}
\hypersetup{urlcolor=UrlColor}
\maxdeadcycles=1000
\allowdisplaybreaks
\textwidth 7.5 in 
\hoffset -1 cm 
\newcommand{\comment}[1]{\textcolor{blue}{\textit{#1}}}
\newcommand{\ashok}[1]{\textcolor{red}{\textit{Ashok:#1}}}
\newcommand{\sean}[1]{\textcolor{cyan}{\textit{Sean:#1}}}

@@

\title{Spin effect on Christodoulou memory for merging blackhole binaries}
\author{Ashok Choudhary}\email{aschoudhary@mix.wvu.edu}
\author{Sean T. McWilliams}\email{sean.mcwilliams@mail.wvu.edu}
\affiliation{Department of Physics and Astronomy, West Virginia University, Morgantown, WV 26506, USA}

\begin{abstract}
	The Christodoulou memory, which is a nonlinear memory effect sourced by the gravitational wave stress tensor, produce a growing, nonoscillatory change in the gravitational wave "plus" polarization. This results in the permanent displacement of a pair of freely falling test masses after the gravitational wave has passed. While the memory contribution during early in-spiral is well described by Post-Newtonian approximation, the contribution at merger must be obtained from Numerical relativity simulations. The Post-Newtonian corrections upto 3PN order to the gravitational wave memory for quasicircular, inspiralling compact binaries for non spinning has been computed by Favata et al.[Physical Review D 80, 024002 (2009)]. Here we include the Spin contribution at leading order and calculate the memory effect upto 3PN order. We also compute memory contribution at merger from Numerical relativity using Weyl curvature component $\psi_4$ after removing the unphysical low frequency contribution due to spectral leakage. We use the method described in [Class. Quantum Grav. 28 195015] for integration which control the amplification of the unphysical frequency resulting from spectral leakage. 
\end{abstract}

@@

\section{Introduction}
Gravitational wave memory effect leads to a permanent displacement of test masses after the gravitational wave has passed through it. Nonlinear memory effect also called Christodoulou memory, arises from a change in the radiative multiple moment that is sourced by the energy flux of the radiated gravitational wave. It can be understood in the following way. Consider the Einstein field equation in harmonic gauge. [add ref] 
\begin{subequations}
\begin{align}	
\Box \bar{h}^{\alpha \beta} & = -16\pi (-g)(T^{\alpha\beta} + t_{LL}^{\alpha\beta})-\bar{h}^{\alpha\mu}_{,\nu}\bar{h}^{\beta\nu}_{,\mu}+\bar{h}^{\mu\nu}\bar{h}^{\alpha\beta}_{,\mu\nu}\\
\bar{h}^{\alpha \beta},_{\beta} & = 0
\end{align}
\end{subequations}
where
\begin{equation}
\bar{h}^{\alpha \beta} = \eta^{\alpha\beta}-\sqrt{-g}g^{\alpha\beta} 
\end{equation}
is the gravitational field tensor, $g$ is the determinant of the metric $g_{\alpha\beta}$, $T^{\alpha\beta}$ is the matter stress-energy tensor, $t^{\alpha\beta}_{LL}$ is the Landau-Lifshitz pseudotensor, $\Box \equiv  -\partial^{2}_{t} + \nabla^{2}$ is the flat-space wave operator, a comma denotes a partial derivative $(,_{\mu}\equiv \partial_{\mu})$, and $\nabla^{2}$ is a flat-space Laplace operator. The Landau-Lifshitz term is the Gravitational wave stress tensor[add ref].
\begin{equation}
	T^{gw}_{jk}=\frac{1}{32\pi}\langle h^{TT}_{ab,j}h^{TT}_{ab,k}\rangle\approx T^{gw}_{00}n_{j}n_{k} = \frac{1}{R^{2}}\frac{dE^{gw}}{dt d\Omega}n_{j}n_{k}
\end{equation}
where $\frac{dE^{gw}}{dtd\Omega}$ is the GW energy flux, $n_{j}$ is a unit radial vector, the angle-bracket mean to average over several wavelengths, $h^{TT}_{ab,j}=\bar{h}^{TT}_{ab,j}$. When we apply standard Green's function to right hand side of equation (1.1a), we obtain the following correction term to GW field [add ref]
\begin{equation}
	\delta h^{TT}_{jk} = \frac{4}{R}\int_{-\infty}^{T_{R}} dt'\Bigg[ \frac{dE^{gw}}{dt'd\Omega'}\frac{n'_{j}n'_{k}}{(1-\boldsymbol{n'.N})}d\Omega'\Bigg]^{TT}
\end{equation}
where $T_{R}$ is the retarded time. This equation shows that part of the distant GW is sourced by loss of GW energy.

\section{POST-NEWTONIAN WAVEFORM POLARIZATION AND MODE DECOMPOSITION}
Following Kidder [add ref] we introduce a asymptotically-flat radiative coordinate system described by spherical coordinates $(T, R, \Theta, \Phi)$ and their orthogonal basis vectors $(\vec{e}_T, \vec{e}_R, \vec{e}_{\Theta}, \vec{e}_{\Phi})$. The coordinate origin is center-of-mass of the source. The retarded time in radiative coordinates is $T_R = T - R$
\\
The mode decomposition of gravitational wave polarization is given by
\begin{equation}
	h_+ - \mathit{i}h_{\times} = \sum_{l=2}^{\infty}\sum_{m=-l}^{m=l}h^{lm} \,  _{-2}Y^{lm}(\Theta, \Phi)
\end{equation}
where
\begin{equation}
	h^{lm}= \frac{G}{\sqrt{2}\, R \, c^{l+2}}\left[U^{lm}(T_R)-\frac{\mathit{i}}{c}V^{lm}(T_R)\right]
\end{equation}

Here $U^{lm}(T_R)$ and $V^{lm}(T_R)$ are "scalar" mass and current multipoles respectively. The spin weighted spherical harmonics are defined in terms of the Wigner $d$ function by
\begin{equation}
	_{-2}Y^{lm}(\Theta, \Phi) = (-1)^2\sqrt{\frac{2l + 1}{4\pi}}d^{l}_{ms}(\Theta)e^{\mathit{i}m\Phi}
\end{equation} 
Here
\begin{align}
	d^{l}_{ms}&=\sqrt{(l+m)!(l-m)!(l+s)!(l-s)!}\\
	&\times \sum_{k=k_i}^{k_f}\frac{(-1)^k(\sin\frac{\Theta}{2})^{2k+s-1}(\cos\frac{\Theta}{2})^{1l+m-s-2k}}{k!(l+m-k)!(l-s-k)!(s-m+k)!}
\end{align}
where $k_i$ = max$(0, m-s)$ and $k_f$ = min$(l+m, l-s)$.

\section{MEMORY CONTRIBUTION TO THE RADIATIVE-MASS MULTIPOLES}
The GW polarization can be expanded on the basis of spin-weighted spherical harmonics.
\begin{equation}
	\sqrt{2}R(h_{+} - ih_{\times}) = \sum_{l=2}^{\infty}\sum_{m=-l}^{l}(U_{lm} - \mathit{i}V_{lm})_2Y^{lm}(\Theta, \Phi)
\end{equation}
Here $(R,\Theta,\Phi)$ are the spherical coordinates pointing from the source to the observer. The coefficient $U_{lm}$ and $V_{lm}$ are mass and current multipole moments. \\
The memory contribution to the waveform polarization can be evaluated in the the following ways as described in [add referece]

\subsection{Computing angular integral}

\begin{equation}
	\int d\Omega n_{L},
\end{equation}

\begin{equation}
	\int d\Omega _{-2}Y_{l'm'}(\theta,\phi)_{-2}Y_{l''m''}^{*}(\theta, \phi)Y_{lm}^{*}(\theta, \phi)
\end{equation}

\begin{align}
	U_{lm}^{(mem)(1)}=&R^{2}\sqrt{\frac{2(l-2)!}{(l+2)!}}\sum_{l'=2}^{\infty}\sum_{l''=2}^{\infty}\sum_{m'=-l'}^{l'}\sum_{m''=-l''}^{l''}(-1)^{m+m'}\\
	&\times\Bigg\langle \dot{h}_{l'm'}\dot{h}_{l'm'}^{*}\Bigg\rangle G^{2-20}_{l'm'lm'-m''-m}
\end{align}

\subsection{Time derivatives of the memory multipole moments}

\begin{equation}
	\int_{-\infty}^{T_{R}} x^{n}dt=\int_{-\infty}^{T_{R}}\frac{x^{n}}{\dot{x}}dx
\end{equation}

\subsection{Computing time integral over the past history of the source}
sesadsd

% 
\begin{widetext}
\begin{align}\label{eq:FbydE}
\frac{dx}{dt} & = \frac{64\, x^5  \, \eta}{5} \left\{1 + x \left[-\frac{743}{336}-\frac{11 \eta }{4} \right] 
    + x^{3/2} \left[4 \pi-\frac{113}{12}\delta \, \chiadL- \left(-\frac{113}{19} + \frac{19 \, \eta}{3}  \right)\chisdL \right] \right. \nonumber \\ 
    &\quad + x^2 \left[ \frac{34103}{18144} + \frac{13661 \, \eta}{2016} + \frac{59 \, \eta^2}{18} + \left( \frac{719}{96}- 30 \eta\right)(\chiadL)^2 +\chiaSqr \left(-\frac{233}{96}+10 \eta \right) \right. \nonumber\\
   &\qquad+\frac{719 \chiadL \, \chisdL \delta
   }{48}-\frac{233 \chisDchia \delta}{48} \left. + \, (\chisdL)^2 \left(-\frac{\eta }{24}-\frac{719}{96}\right)+\chisSqr \left(\frac{7 \eta
   }{24}+\frac{233}{96}\right) \right] \nonumber \\ 
    &\quad + x^{5/2} \left[-\frac{4159 \, \pi}{672}-\frac{189 \, \pi \, \eta}{8} + \chiadL \delta  \left(-\frac{31319}{1008} + \frac{1159 \, \eta}{24}\right)+\chisdL \left(-\frac{31319}{1008} + \frac{22975 \, \eta}{252} -\frac{79 \, \eta^2}{3}\right) \right]  \nonumber\\ 
    &\quad \left. + x^3 \left[\frac{16447322263}{139708800} + \frac{16 \, \pi^2}{3} -\frac{1712 \, \gamma_{E}}{105} -\frac{56198689 \, \eta}{217728} + \frac{451 \, \pi^2 \, \eta}{48} + \frac{541 \, \eta^2}{896} -\frac{5605 \, \eta^{3}}{2592} -\frac{80 \,\pi \delta\chiadL }{3} \right.\right. \nonumber \\
    &\qquad \left.\left. + \left(\frac{575}{448} + \frac{565 \, \delta^2}{9}-\frac{65815 \, \eta}{4032} + \frac{89 \, \eta^2}{2}\right)(\chiadL)^2 + \left(-\frac{145}{448}+\frac{21985 \,\eta}{4032} -\frac{89 \, \eta^2}{6}\right)\chiaSqr + \right(-\frac{80 \, \pi}{3} + \frac{40 \, \pi \, \eta}{3} \right. \nonumber \\
    &\qquad  \left.+\delta\left(\frac{258295}{2016} -\frac{9203 \, \eta}{96}\right)\chiadL\left)\chisdL + \left(\frac{258295}{4032} -\frac{48773 \, \eta}{576} + \frac{3041 \, \eta^2}{144}\right)(\chisdL)^2 + \delta\left(-\frac{145}{224} + \frac{2143 \, \eta}{288}\right)\chisDchia \right.\right.\nonumber \\ 
    &\qquad\left.\left.+\left(-\frac{145}{448} + \frac{1891 \, \eta}{576} -\frac{7 \, \eta^2}{144}\right)\chisSqr - \frac{3424 \, \log[2]}{105} - \frac{1712 \log[x]}{210}\right]\right\}
\end{align}
\end{widetext}

\subsection{RESULTS: MEMORY CONTRIBUTION TO THE POST-NEWTONIAN WAVEFORM OF QUASI-CIRCULAR, INSPIRALLING BINARIES}
The memory contribution to the spin-weighted spherical-harmonic modes of the polarization waveform [ref eqn]. These quantities are related via 
\begin{equation}
	h_{l0}^{(mem)} = \frac{\alpha}{\sqrt{2} \, R} U_{l0}^{(mem)} = 8 \sqrt{\frac{\pi}{5}}\frac{\eta M x}{R} 
\end{equation}
where we have followed the notation of Sec. 9 of Ref [add referece]. The notational parameter $\alpha$ accounts for the two commonly used choices for the polarization triad 
\begin{equation}
	content...
\end{equation}
(\ashok{Need to add}).\\
The results of polarization modes in terms of $\hat{H}_{l0}$ are:
\begin{widetext}
\begin{align}\label{eq:H20}
\hat{H}_{20} &= \alpha \frac{5}{15\sqrt{6}}\left\{ 1 + x\left(-\frac{4075}{4032} + \frac{67 \, \eta}{48}\right)+ x^{3/2}\left(\left(-\frac{7}{1440} -\frac{27 \, \delta}{10} + \frac{7 \,\eta}{360}\right)\chiadL+\left(-\frac{27}{10} -\frac{7 \, \delta^3}{1440} + \frac{22 \, \eta}{15}\right)\chisdL\right)\right. \nonumber \\
& \quad \left. x^2\left(-\frac{151877213}{67060224} -\frac{123815 \, \eta}{44352} +\frac{205 \, \eta^2}{352} -\left(\frac{1459}{864} -\frac{1441 \, \eta}{216}\right)\chiaSqr - \left(\frac{27 \, \delta}{8} +\frac{\delta^3}{432}\right)\chisDchia - \left(\frac{27}{16}+\frac{\delta^4}{864} -\frac{\eta}{12}\right)\chisSqr\right)\right. \nonumber \\
& \quad \left. x^{5/2}\left(-\frac{253 \, \pi}{336}+ \frac{253 \, \pi \, \eta}{84} -\left(-\frac{631}{96768} + \frac{2729857 \, \delta}{84672} -\frac{146597 \, \pi \, \delta}{1764}+\frac{5 \, \eta}{378}+\frac{3007 \, \delta \, \eta}{1008} -4\pi \, \delta \, \eta +\frac{311 \, \eta^2}{6048}\right)\chiadL\right.\right. \nonumber\\
& \qquad\left. -\left(\frac{2729857}{84672}-\frac{146597 \, \pi}{1764}-\frac{631 \, \delta^3 }{96798} -\frac{44111 \, \eta}{1764}+\frac{4852 \, \pi \, \eta}{63}-\frac{311 \, \delta^3 \, \eta}{24192}-\frac{1121 \, \eta^2}{252} +\frac{68 \, \pi \, \eta^2}{7}\right)\chisdL\right) \nonumber \\
& \quad  x^3\left[-\frac{4397711103307}{532580106240} + \left( \frac{700464542023}{13948526592} - \frac{205 \, \pi^2}{96}\right)\eta + \frac{69527951 \, \eta^2}{166053888} + \frac{1321981 \, \eta^3}{5930496} -  \left(-\frac{160867}{13824} -\frac{791 \, \delta}{27648} + \right.\right. \nonumber \\
& \qquad \left. \frac{7345 \, \delta^2}{576} + \left(\frac{3953767}{96768}+\frac{791 \, \delta}{6912}\right)\eta + \frac{22885 \, \eta^2}{1152}\right)\chiaSqr -\left(-\frac{275 \, \pi}{12} + \frac{7 \, \pi \, \delta^3}{768} + 17 \, \pi \, \eta \right)\chisdL + \left(\frac{47869}{43008}  \right.\nonumber\\
& \qquad \left.\left. - \frac{791 \, \delta^3}{27648} + \frac{743 \, \delta^4}{387072} + \left(-\frac{410453}{16128} + \frac{133 \, \delta^3}{6912} + \frac{11 \, \delta^4}{4608}\right)\eta + \frac{10229 \, \eta^2}{1152}\right)\chisSqr - \left( \frac{7 \, \pi}{768}-\frac{275 \, \pi \, \delta}{12}-\frac{7 \, \pi \, \eta}{192}\right.\right. \nonumber \\
&\qquad \left.\left.\left. + \left(-\frac{791}{27648} + \frac{47869 \, \delta}{21504} + \frac{743 \, \delta^3}{193536} - \frac{791 \, \delta^4}{27648} + \left(\frac{77}{576} - \frac{71749 \, \delta}{2304} + \frac{11 \, \delta^3}{2304}\right)\eta - \frac{133 \, \eta^2}{1728}\right)\chisdL\right)\chisdL
\right]
\right\}
\end{align}

\end{widetext}
